---
title: The 2016 NBA Draft Class Was Historically Bad
author: Michael Groesbeck
date: '2017-05-11'
slug: this-nba-draft-class-was-bad
categories:
  - R
tags:
  - NBA
  - R
description: ''
featured: ''
featuredalt: ''
featuredpath: ''
linktitle: ''
---


```{r, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warnings=FALSE)
```


On a semi-recent episode of [The Lowe Post](add_url.com), Jeff Van Gundy, talking about his choice for Rookie of the Year, asked- "Is this the worst [rookie class] as far as initial contribution? Has there ever been less contribution to winning?" 

To answer these questions, I looked at each draft class since 1989 (when the draft was changed to two rounds) and their performance in that year's season. To be clear, this is only looking at immediate contribution, and not long-term success. Only those who played the same year they were drafted are being considered- so players like Joel Embiid, who did not play their first year, are not in the data.

## The Data

The data is from [basketball-reference.com](www.basketball-reference.com). First, a year's [draft class](http://www.basketball-reference.com/draft/NBA_2016.html) was pulled, then this was joined with that same season's [player statistics](http://www.basketball-reference.com/leagues/NBA_2017_advanced.html). This was done for every year 1989 - 2016. The following script was used to scrape and prepare the data for analysis. 

```{r, eval=FALSE}
library(rvest)
library(tidyverse)
library(stringr)

# function used to pull a single years data
get_year_data <- function(year){
  
  draft_url <- paste0("http://www.basketball-reference.com/draft/NBA_", 
                      year, ".html")
  
  # return just the player name and pick
  draft <- read_html(draft_url) %>% 
    html_node("table") %>% 
    html_table(trim=TRUE, header=FALSE) %>% 
    filter(!is.na(as.numeric(X1))) %>% 
    select(Pick = X1, 
           Player = X4)
  
  # 2016-17 season is under 2017 url so add one to year
  stat_year <- year + 1
  
  adv_stats_url <- paste0("http://www.basketball-reference.com/leagues/NBA_", 
                      stat_year, "_advanced.html")
  
  adv_stats <- read_html(adv_stats_url) %>% 
    html_node("table") %>% 
    html_table(trim=TRUE, header=TRUE)
  
  # remove asterisk from players name (indicates all-star?)
  adv_stats$Player <- str_replace_all(adv_stats$Player, "\\*", "")  
  
  tot_stats_url <- paste0("http://www.basketball-reference.com/leagues/NBA_", 
                      stat_year, "_totals.html")
  
  tot_stats <- read_html(tot_stats_url) %>% 
    html_node("table") %>% 
    html_table(trim=TRUE, header=TRUE) %>% 
    select(Player, TRB, AST, PTS) %>%
    mutate(Player = str_replace_all(Player, "\\*", "")) %>% 
    filter(!duplicated(Player))
  
  # get year, number of teams, and total minutes played that season
  year_info <- data_frame(
    Year = year,
    Teams = length(unique(adv_stats$Tm[!adv_stats$Tm %in% c("TOT", "Tm")])),
    Minutes = sum(as.numeric(adv_stats$MP[adv_stats$Tm != "TOT"]), na.rm=TRUE),
    WinShares = sum(as.numeric(adv_stats$WS[adv_stats$Tm != "TOT"]), na.rm=TRUE),
    Points = sum(as.numeric(tot_stats$PTS), na.rm=TRUE),
    Rebounds = sum(as.numeric(tot_stats$TRB), na.rm=TRUE),
    Assits = sum(as.numeric(AST), na.rm=TRUE))
  
  inner_join(draft, adv_stats, by = "Player") %>% 
    select(Pick, Player, Tm, Pos, G, MP, OWS, DWS, WS) %>% 
    inner_join(tot_stats, by = "Player") %>% 
    mutate(
      Pick = as.numeric(Pick), 
      Player, 
      Pos, 
      G = as.numeric(G), 
      TRB = as.numeric(TRB),
      AST = as.numeric(AST),
      PTS = as.numeric(PTS),
      MP = as.numeric(MP), 
      OWS = as.numeric(OWS), 
      DWS = as.numeric(DWS), 
      WS = as.numeric(WS),
      Year = year) %>% 
    # removes the splits for players traded mid-year
    filter(!duplicated(Player)) %>% 
    left_join(year_info, by="Year")
}

# get data from 1989 - 2016
nba <- lapply(1989:2016, get_year_data) %>% 
  bind_rows() %>% 
  # in 2015 a 2nd marcus thornton was drafted who did not play the next year
  filter(Player != "Marcus Thornton" | Year != 2015) %>% 
  # in 2007 a 2nd marcus williams was drafted, this lists the others stats hard fix
  filter(Player != "Marcus Williams" | Year != 2007) %>% 
  rbind(c(33, "Marcus Williams", "TOT", "PF", 11, 36, -0.1, 0, -0.1, 12, 3, 10, 2007, 30, 594122))

```

The data results can be found [here](https://github.com/). 

## Analysis and Results

The first thing that can be used as a measure of contribution is minutes played. In order to compare across seasons, we compute what percentage of minutes played in a season were by rookies.

```{r, echo=FALSE, fig.width=8}
library(tidyverse)
library(scales)
library(ggthemes)

nba <- read_csv("~/Dropbox/projects/nba_draft_usage.csv")

by_year <- nba %>% 
  group_by(Year, Minutes) %>% 
  summarize(MP = sum(MP, na.rm=TRUE),
            WS = sum(WS, na.rm=TRUE)) %>% 
  mutate(Perc_MP = MP / Minutes,
         Perc_WS = WS / Minutes)

theme_rom <- theme_fivethirtyeight() + 
  theme(axis.text.x=element_text(angle=90, vjust=.4),
        legend.position="bottom", axis.title=element_text(),
        panel.background = element_rect(fill="#fffefc"),
        plot.background = element_rect(fill="#fffefc"),
        legend.background = element_rect(fill="#fffefc"),
        legend.key = element_rect(fill="#fffefc"),
        panel.grid.major = element_line(color="#d2d2d8"))

ggplot(by_year, aes(Year, Perc_MP)) +
  geom_line(color="#00369C", size=1.5) +
  scale_x_continuous(breaks = 1989:2016) +
  scale_y_continuous(labels = percent) +
  theme_rom +
  labs(title="Contribution by NBA Draft Classes", subtitle="Minutes Played",
       y="Percent of total minutes") +
  theme(axis.title.x = element_blank())

ggplot(by_year, aes(Year, Perc_WS)) +
  geom_line(color="#00369C", size=1.5) +
  scale_x_continuous(breaks = 1989:2016) +
  theme_rom +
  labs(title="Contribution by NBA Draft Classes", subtitle="Win Shares",
       y="Win Shares / Total Minutes") +
  theme(axis.title.x = element_blank())

```

*test test*

